### 2016-10-03 Paper_Final.tex
Final editing pass.

\noindent\Large \textsc{Aaron Rudkin\footnote{I thank Natalie McNaught from the New Zealand Electoral Commission for her assistance securing and digitizing paper election records. My thanks also to Kathy Bawn, Adam Boche, Jeff Lewis, Luke Sonnet, and Michael F. Thies for input at various stages during this project. Any errors or omissions are my own. Replication data and code is available at: \textcolor{blue}{https://github.com/aaronrudkin/CabinetEffect}}}
\vspace*{0.1\textheight}

@@

\textmd{Although a long body of research has considered cabinet formation from the perspective of parties and coalitions, research on individual incentives to join cabinet has been limited thus far. Electoral benefits, which may stem from particularistic spending, differential credit claiming capability, or party investment, provide one motivation to join cabinet. Martin (2016) uses Irish data to present evidence for electoral returns from cabinet membership. I extend Martin's analysis using a novel electoral dataset I constructed, examining Canada and New Zealand post-1945. In the Canadian data, I find a modest but substantively relevant positive effect, robust to alternate specifications and aggregations of the data. By contrast, the New Zealand data shows little to no effect. Limitations including endogeneity in cabinet selection, data limitations, and the failure of the key result to replicate across institutional and cultural contexts suggest caution is warranted before drawing broader global conclusions.}

@@

\section*{Introduction}
Cabinet formation has long been considered from the perspective of parties and coalitions. When forming a cabinet, parties must decide who to include to please constituencies, geographies, factions within the party, and coalition partners. But parties are not monoliths, they are aggregations of members with individual preferences and goals. (M{\"u}ller and Str{\o}m 2000) Among the goals sought by any legislator are ability to achieve policy outcomes; personal and career satisfaction; service to party; and most importantly re-election. (Fenno 1978) Joining cabinet plausibly assists with the first three of these goals, but \textit{do cabinet ministers receive an electoral return in exchange for joining cabinet}? The answer is relevant to MPs deciding joining a cabinet, parties considering electoral strategy, and researchers seeking to understand the behaviour of legislators and the incentives they work with.

@@

Downs (1957), Key (1964), and Mayhew (1974) are among early authors who characterized legislative and party behaviour as being animated by desire for re-election. Mayhew offers three mechanisms through which legislators earn re-election: advertising, credit claiming, and position taking. Fenno (1978) takes a broader view of legislator behaviour and expands on the ways legislators form relationships with their constituents to secure re-election. An essential tension exists: on one hand, incumbent governments pay an electoral cost for governing. (Mershon 1996) On the other hand, a widely documented incumbent advantage makes staying elected an easier proposition than winning an open election. Incumbents deter competition and enjoy resource advantages as compared to their competition. (Erikson 1971; Carson 2005; Ban, Llaudet, and Snyder 2016; Gelman and King 1990)

Legislators seek to maximize their incumbency advantage while minimizing their costs of governing. Cain, Ferejohn, and Fiorina (1984) explain that constituency service can be used to build a ``personal vote''. The incentives to earn a personal vote vary by electoral and institutional context. Electoral systems such as single-member plurality where voters choose candidates directly offer the highest incentive for a personal vote, because members can be rewarded or protected by developing a reputation beyond that of their party. (Carey and Shugart 1995) The tools available to members also vary by institutional context. Parliamentary systems impose whipped voting and use vote scheduling rules to prohibit backbenchers from advancing substantive business, theoretically limiting the ability of members to distinguish themselves from the party. This has ramifications for the way in which legislators see their roles, what they hope to accomplish in a legislature, the decision to continue serving, and how they situate themselves with respect to their voters.

Studies of particularistic legislative spending find that legislators gain electoral advantage through the delivery of spending and targeted policy goods. In a US congressional context, research focuses on the ability of individual legislators to secure spending for their home districts. Parliamentary systems limit the ability of backbenchers to achieve these benefits on their own. Instead, parties act as a cartel to protect their members. In one telling of this story, the governing party uses distribution of particularistic benefits to benefit all its members. But in Martin's telling of the story, cabinet ministers have control over the shape of policy and spending within their particular ministry and use this control in concert as a cartel to support each other through the targeted provision of these goods: what he terms ``executive particularism.'' (Martin 2016) Preliminary work in a variety of national settings finds that such particularistic spending does occur at the cabinet level. (John, Ward, and Dowding 2004; Denemark 2000)

@@

In addition to executive particularism, exposure may also be a driver of electoral success. Additional media exposure and the opportunity to hold events in concert with spending or legislation may assist cabinet ministers in credit claiming. Evidence indicates that credit claiming is an important part of the ability to capitalize electorally on the delivery of spending or policy goods. (Grimmer, Messing, and Westwood 2012; Bickers, Evans, Stein, and Wrinkle 2007) To the extent that ministers enjoy a differential ability to claim credit, they should reap rewards for doing so.

\section*{Case Selection}
The ideal cases to test for the cabinet effect are Westminster parliamentary democracies with competitive multi-party elections, incumbents elected primarily from district-level elections, a norm against district-switching by incumbents, and a tendency towards single-party government. Selecting for the latter allows the research to avoid engaging the complexities of cabinet appointment in a coalition setting and intra-election coalition changes. (Laver and Shepsle 1996) I narrow my choice of cases to Australia, Canada, Ireland (studied by Martin), New Zealand, and the United Kingdom and selected Canada and New Zealand from these initial cases based on scope limitations.

Canada is a Westminster parliamentary democracy established in 1867. It features a 338 member lower house, an appointed upper house that as a matter of convention does not impede government business, and an SMP electoral system. Canada has never had a coalition government, preferring minority governments by convention. Canada has historically been governed by one of two parties: the Liberal Party and the Progressive Conservative Party (after 2004, the Conservative Party). Minor regional parties have achieved some electoral success. The effective number of political parties\footnote{A measure of party competition, where $N = \frac{1}{\sum_{i=1}^n p_i^2}$, with parties 1 through \textit{n} each occupying a $p_i$ share of seats. The measure was developed by Laakso and Taagepara (1979).} during the time studied varies from 1.5 to 3.2. Once candidates have contested an election, they do not switch ridings in subsequent elections, although parachuting (carpet-bagging) of star candidates into politically opportune ridings occurs among first-time candidates. (Koop and Bittner 2011) Elections in Canada are relatively low cost with bans on corporate and union donations, extensive public funding, and spending caps. Major party candidates typically spend in the order of \$75,000 per four-year election.

Canadian Prime Ministers enjoy broad latitude to make cabinet appointments, subject only to norms with respect to the representation of each province in cabinet, the allocation of certain portfolios to relevant geographic representatives (for example, the Fisheries ministry is traditionally given to an MP from one of Canada's coastal provinces), and increasingly diversity of gender and ethnicity. A survival analysis of cabinet entry in Canada found that gender, legal experience, education, age, and past ministerial tenure were major predictors of selection. (Kerby 2009) Over time, cabinets in Canada have become increasingly professionalized and technocratic, a trend also observed in other developed democracies (Lammers and Nyomarkay 1982; Pekkanen, Nyblade, and Krauss 2013; Berlinski, Dewan, Dowding, and Subrahmanyam 2009)

New Zealand is a Westminster parliamentary democracy established in 1852. Historically it featured a lower house with fewer than 100 members, a perfunctory upper house abolished in 1950, and an SMP electoral system with reserved electorates for voters who self-identified as indigenous Maori. Dissatisfaction with the two major political parties, National and Labour, in part because of repeated majority governments elected by weak pluralities of popular vote support, led New Zealand to enact electoral reform beginning with the 1996 national election. The current electoral system is closed-list mixed-member proportional with a low minimum vote quotas, and consists of approximately 121 seats including 71 electorate seats. The effective number of political parties varied from 1.76 to 2.16 before MMP and from 2.78 to 3.76 after MMP. Each government post-MMP has featured either supply agreements or coalitions with minor parties. Elections in New Zealand are extremely low-expense, with major party candidates spending in the order of \$20,000 per three-year election.

In principle, Labour governments in New Zealand allow their entire caucus to vote on cabinet membership, but in practice Prime Ministers are allowed considerable leeway to expand or shrink the size of cabinet and allocate portfolios among those selected. (Alley 1989; Wood 1989) Cabinets typically include representation from both of New Zealand's islands and from urban and rural areas. The Minister of Maori Affairs position is often filled by a Maori legislator, and increasingly the representation of women in cabinet has been a relevant factor. (Curtin 2015)

Both countries share a variety of cultural and institutional factors: both feature high levels of parliamentary discipline and practice cabinet collective and individual responsibility. Expulsion from cabinet, caucus, and denial of renomination are the primary punishments available to enforce party discipline. Neither country has a formal primary election mechanism, instead delegating candidate nominations to district-level parties, who by custom do not remove incumbents for reasons other than non-compliance with party leadership or scandal. (Levine 1979; Wood 1989) Finally, both cases are known to practice ``government by cabinet'', with inert backbenches and powerful cabinets. (Henderson 1989)

@@

\section*{Data}
The dataset I make use of is novel. Existing election datasets suffer from limitations that render them inappropriate for use here: Much of the data is aggregated above the constituency level; constituency level data is often missing names and characteristics of individual candidates necessary to make the data a panel; districts and candidates names vary across elections; and data typically does not include cabinet membership covariates.

As a result, I assembled a dataset containing full, constituency-level electoral data for all candidates and legislators in Canada from 1867 forward, and New Zealand from 1945 forward.\footnote{Canadian cabinet ministers before 1931 were expected to immediately resign from the House of Commons upon being appointed and then run in by-elections to regain the seat they resigned. To avoid the complexity of this feature of the data and the tail end of a party system transition occurring in the 1930s, I discard data before the 1945 federal election.} Canadian data was obtained by writing a series of web scrapers and harvested from Canadian Library of Parliament sources. New Zealand data was assembled using a combination of previously published volumes\footnote{Wood 1996; Norton 1988}, paper returns supplied by the NZ Electoral Commission and digitized for the first time for this project, and web scrapers for data from 1996 onwards.

After combining and cleaning the data, the next challenge was being able to connect legislators across elections. To deal with internal inconsistencies in the reporting format of district names, party names, and candidate names, I wrote a text analysis program that uses candidate metadata to match candidates across elections to assemble the panel. Clear candidate matches were automatically identified, while unclear matches prompt the user to make a manual coding decision which is saved for replication.

My dataset is exhaustive as to candidates who stood for election in a district, but a small number of New Zealand MPs have been elected from the party list without contesting an electorate, including some cabinet ministers. These MPs are not represented in the dataset as they did not stand for election individually.\footnote{One example: Deputy Prime Minister Bill English, a long-time electorate incumbent who switched to list-only despite huge electorate majorities to minimize his travel commitment and time away from family. There is no electoral benefit to being list-only.}

@@

The primary limitation when testing this hypothesis is severe endogeneity. The ``gold standard'' to tackle this problem would be to experimentally appoint cabinet ministers through random assignment and measure causal effects. Reality is not so kind: Parties select cabinet ministers to satisfy goals including rewarding key constituencies, demographics, and policy demanders within the party and coalition partners. Moreover, parties have access to private signals about incumbent quality learned during the recruitment, vetting, and campaigning process that are not fully accounted for in vote share at the time of their first election. These latent candidate qualities may become obvious during the process of legislating and materialize in future results, but are not strictly speaking a cabinet effect: Koop and Bittner (2011) argue that ``star candidates'' are much more likely to be appointed to cabinet. In light of these concerns, the claims made here cannot be said to be causal: effects in this paper are associations, suggestive of a relationship, but not decisive.

@@

\setstretch{1.5}
\begin{itemize}
\item \textit{VotePct}: The vote share of the candidate.
\item \textit{CabinetNow}: The candidate was a member of the cabinet during the preceding term. I exclude ministers outside cabinet, junior ministers, ministers without portfolio, and other quasi-cabinet roles but include deputy Prime Ministers and Attorneys General.\footnote{I also do not model ministerial exit from cabinet or role-switching mid-term, so in principle some of the \textit{CabinetNow} incumbents have been removed from cabinet before the election. Evidence suggests that ministerial exit is predicted by poor performance (Berlinkski, Dewan, and Dowding 2010) or scandal, (Dewan and Myatt 2007; Dewan and Dowding 2005) so this modeling decision should cause effect estimates to be biased conservatively as poor ministers drag the overall effect down.}
\item \textit{CabinetImportant}: The candidate was an ``important'' minister of the cabinet during the preceding term. ``Importance'' is subjective and here includes Defence, Health, Justice, Finance, Public Works, Revenue, and Foreign ministries. This variable exists to capture heterogeneity in the cabinet effect; senior ministers are presumed to have more discretion with respect to particularistic spending, more control over policy outcomes, and greater capacity to cultivate a personal vote through media exposure and credit claiming.
\item \textit{CabinetPM}: Dummy variable for the Prime Minister, regardless of whether they were an incumbent member of the legislature. The inclusion of a control for Prime Ministers is intended to absorb the positive electoral effects expected for a sitting Prime Minister to avoid their misattribution to cabinet membership.
\item \textit{Incumbent}: The candidate is an incumbent member of the legislature. I treat candidates as incumbent even if redistricting occurs, or if the candidate was previously elected by party list and now contests an electorate. Incumbency is known to be a a strong predictor of re-election.
\item \textit{TermsServed}: The total unbroken number of terms served by a candidate. Tenure does not reset if an incumbent resigns and runs in a by-election, but does reset if a candidate loses or resigns and later returns to the legislature. I treat by-elected members as serving a full time. Tenure is present in the model to capture the accumulated effects of developing a personal vote (or, at least, a reputation of competency) in an incumbent's district over time.
\item \textit{Age}: The age of the candidate at the time of the election. Only available for a subset of Canadian candidates who were elected to office at least once. Age was gathered to allow for a proxy for signals of candidate quality including occupation, past electoral and career experience.
\item \textit{Elected}: Binary indicator for election outcome.
\item \textit{Party}: Party label the member stood for election under. I recode minor parties or splinter factions to be independent.
\item \textit{PartyInGovt}: Is the candidate running for the party currently governing? In New Zealand post-1996, I include junior coalition partner parties but not parties in supply agreements with the governing party. This variable was constructed in order to account for the ``electoral costs of governing''.
\end{itemize}
\setstretch{2}

@@

The effect's appearance in Canada contrasts its absence in New Zealand. Effect estimates in the main model are small (1.3\%) with confidence intervals spanning substantively unimportant territory. Martin's model reveals a small effect of 2.1\%, but is likely picking up correlation between tenure and cabinet selection due to the excluded tenure variable. Additional specifications are inconsistent as to whether the effect size is substantively interesting. Disaggregating by decade reveals no consistency. The relatively parsimonious models have high goodness of fit, indicating that much of the variation in performance is driven by the covariates modeled, particularly party label, despite limited cabinet effects. Further analysis as to why New Zealand may behave differently than Ireland or Canada is left until after presentation of results.

@@

Cabinet ministers are more popular, but not necessarily because they are cabinet ministers. In fact, there is evidence the causality is exactly reversed: A cross-sectional logistic regression of probability being selected into the Canadian Cabinet based on age, tenure, and previous performance suggests that previous performance matters a great deal. In the modal case, a 50 year old sophomore MP of the governing party is much 50\% more likely to be appointed a cabinet minister if she comes from a 70\% blowout win than a 40\% plurality win.

@@

\setstretch{1.25}
\begin{eqnarray*}
	\Delta y_i &=& \beta X_i + \epsilon_i =\\
	y_{i,t} - y_{i,t-1} &=& \beta X_i + \epsilon_i = \\
	y_{i,t} &=& \beta X_i + y_{i,t-1} + \epsilon_i
\end{eqnarray*}
\setstretch{2}
The assumption about the relationship between previous performance and current performance is explicit. Incumbents begin with their previous vote share, and the model fit determines how the covariates are associated with deviations from that share.\footnote{This is conceptually difference from a true ``fixed effects'' model, which also differences out predictor covariates.} A second choice to control for past performance is a lagged dependent variable. This is a weaker assumption; in this model, past performance is merely associated with present vote share, with the strength of the relationship permitted to vary:

@@

\setstretch{2} As expected, endogeneity stemming from the relationship between popularity and cabinet selection in the previous term drives some of the effect. Controlling for past performance, the coefficient estimate of \textit{CabinetNow} (as well as tenure and party) decreases markedly in Canada. In the primary, controlling-for-past-performance model, the effect size is reduced to 1.2\% on average (with the confidence interval permitting interpretations of anything from essentially no effect to an effect slightly smaller than initial cross-sectional estimates). The $\Delta VotePct$ model effects are even smaller still at 0.8\% and include the possibility of no effect.\footnote{A third possibility to control for past performance is to introduce individual fixed effects. This model also results in a smaller estimate of approximately 0.7\% effect for cabinet service with a confidence interval including no effect.} In New Zealand, meanwhile, the effect is negative in all three specifications, although modest in each.

Roughly three quarters of the original dataset are dropped to run these models, causing a severe loss of power---but it is clear that the effect magnitude is reduced in Canada and entirely compromised in New Zealand. Serial autocorrelation in the error structure induced by this strategy is another potential threat to estimate validity. Disaggregation of the CA Control model by decade reveals further inconsistency in the result, with striking effects in the 1980s and 1990s, weaker effects in the 1950s, and near-zero or weakly negative effects in other decades.

@@

Again, the effect is visible and substantively relevant, although modest in size, in Canada. Logistic regression coefficients are not clearly interpretable, but when transformed into odds ratios and predicted probabilities, they become clearer. Using bootstrap resampling to estimate the effect of cabinet membership\footnote{\noindent 1000 samples via nonparametric bootstrap, calculating predicted probabilities for representative case and taking difference to recover effect and CI.}, a two-term incumbent Liberal backbencher in a Liberal government has a 73\% (71\% - 76\%) probability to be re-elected while a cabinet minister with the same characteristics has an 81\% (75\% - 86\%) probability of re-election. A 95\% empirical confidence interval of the effect spans 1.2\%-12.5\%, centered at 7.5\%. As with the above example, the effect is about half the size of an additional term of tenure: modest, but real. The result remains invariant to alternate model specifications tested, including consistent positive effects (some with too little power to rule out no effect) across the disaggregation by decade.

@@

The major assumption underpinning the Cox PH model is proportionality; in other words, that covariates affect the hazard rate in a constant way across the time variable. A test of this assumption using Schoenfeld residuals finds insufficient evidence to reject the null hypothesis (proportional hazards) for the \textit{CabinetNow} variable and validates the appropriateness of the Cox PH model.

>>>>

\noindent\Large \textsc{Aaron Rudkin\footnote{I thank Natalie McNaught with the New Zealand Electoral Commission for her assistance securing and digitizing paper election records. My thanks also to Kathy Bawn, Adam Boche, Jeff Lewis, Luke Sonnet, and Michael F. Thies for input at various stages during this project. Any errors or omissions are my own. Replication data and code is available at: \textcolor{blue}{https://github.com/aaronrudkin/CabinetEffect}}}
\vspace*{0.1\textheight}

@@

\textmd{Although a long body of research has considered cabinet formation from the perspective of parties and coalitions, work on individual incentives to join cabinet has been limited thus far. Electoral benefits, which may stem from particularistic spending, differential credit claiming capability, or party investment, provide a motivation to join cabinet. Martin (2016) uses Irish data to present evidence for electoral returns from cabinet membership. I extend Martin's analysis using a novel electoral dataset I constructed, examining Canada and New Zealand post-1945. In the Canadian data, I find a modest but substantively relevant positive effect, robust to alternate specifications and aggregations of the data. By contrast, New Zealand has limited or no effect. Caveats including endogeneity in cabinet selection, data limitations, and the failure of the key result to replicate across institutional and cultural contexts, suggest caution is warranted before drawing broader global conclusions.}

@@

\section*{Introduction}
Cabinet formation has long been considered from the perspective of parties and coalitions. When forming a cabinet, parties must decide who to include to please constituencies, geographies, factions within the party, and coalition partners. But parties are not monoliths, they are aggregations of members with individual preferences and goals. (M{\"u}ller and Str{\o}m 2000) Among the goals sought by any legislator are: the ability to achieve policy outcomes; personal and career satisfaction; service to party; and most importantly re-election. (Fenno 1978) Joining cabinet plausibly assists with the first three of these goals, but \textit{do cabinet ministers receive an electoral return in exchange for joining cabinet}? The answer is relevant to MPs deciding to join cabinet, parties considering electoral strategy, and researchers seeking to understand the behaviour of legislators and the incentives they face.

@@

Downs (1957), Key (1964), and Mayhew (1974) are among early authors who characterized legislative and party behaviour as being animated by desire for re-election. Mayhew offers three mechanisms through which legislators earn re-election: advertising, credit claiming, and position taking. Fenno (1978) takes a broader view of legislator behaviour and expands on the ways legislators form relationships with their constituents to secure re-election. An essential tension exists: on one hand, incumbent governments pay an electoral cost for governing. (Mershon 1996) On the other hand, a widely documented incumbent advantage exists. Incumbents deter competition and enjoy resource advantages as compared to their competition. (Erikson 1971; Carson 2005; Ban, Llaudet, and Snyder 2016; Gelman and King 1990)

Legislators seek to maximize their incumbency advantage while minimizing their costs of governing. Cain, Ferejohn, and Fiorina (1984) explain that constituency service can be used to build a personal vote. The incentives to earn a personal vote vary by electoral and institutional context. Electoral systems such as single-member plurality where voters choose candidates directly offer the highest incentive for a personal vote, because members can be directly rewarded or protected by developing a reputation beyond that of their party. (Carey and Shugart 1995) The tools available to members also vary by institutional context. Parliamentary systems impose whipped voting and use vote scheduling rules to prohibit backbenchers from advancing substantive business, limiting the ability of members to distinguish themselves from the party. This has ramifications for the way in which legislators understand their roles and situate themselves with respect to their voters.

Studies of particularistic legislative spending find that legislators gain electoral advantage through the delivery of spending and targeted policy goods. In a US congressional context, research focuses on the ability of individual legislators to secure spending for their home districts. Parliamentary systems limit the ability of backbenchers to achieve these benefits on their own. In Martin's telling of the story, cabinet ministers evade these restrictions through control over the shape of policy and spending within their particular ministry. They use this control in concert with other ministers as a cartel to support each other through the targeted provision of goods: what he terms ``executive particularism.'' (Martin 2016) Preliminary work in a variety of national settings finds that such particularistic spending does occur at the cabinet level. (John, Ward, and Dowding 2004; Denemark 2000)

@@

In addition to executive particularism, media exposure and events surrounding legislation may also be a driver of electoral success by assisting cabinet ministers in credit claiming. Evidence indicates that credit claiming is an important part of the ability to capitalize electorally on the delivery of spending or policy goods. (Grimmer, Messing, and Westwood 2012; Bickers, Evans, Stein, and Wrinkle 2007) To the extent that ministers enjoy a differential ability to claim credit, we would expect them to reap electoral rewards for making use of it.

\section*{Case Selection}
The ideal cases to test for the cabinet effect are Westminster parliamentary democracies with competitive multi-party elections, district-based electoral systems, a norm against district-switching by incumbents, and a tendency toward single-party government. Selecting for the latter allows the research to avoid engaging the complexities of cabinet appointment in a coalition setting and intra-election coalition changes. (Laver and Shepsle 1996) I narrowed my choice of cases to Australia, Canada, Ireland (studied by Martin), New Zealand, and the United Kingdom and selected Canada and New Zealand from these initial cases based on scope limitations.

Canada is a Westminster parliamentary democracy established in 1867. It features a 338 member lower house, and an SMP electoral system. Canada has never had a coalition government, preferring minority governments by convention. Canada has historically been governed by one of two parties: the Liberal Party and the Progressive Conservative Party (after 2004, the Conservative Party). Minor regional parties have achieved some electoral success. The effective number of parties\footnote{A measure of party competition, where $N = \frac{1}{\sum_{i=1}^n p_i^2}$, with parties 1 through \textit{n} each occupying a $p_i$ share of seats. The measure was developed by Laakso and Taagepara (1979).} during the time studied varies from 1.5 to 3.2. Once candidates have contested an election, they do not switch ridings in subsequent elections, although parachuting (carpet-bagging) of star candidates into politically opportune ridings occurs. (Koop and Bittner 2011) Elections in Canada are relatively low cost with bans on corporate and union donations, extensive public funding, and spending caps. Major party candidates typically spend in the order of \$75,000 per four-year election.

Canadian Prime Ministers enjoy latitude to make cabinet appointments as they see fit, subject only to norms with respect to the representation of each province in cabinet, the allocation of certain portfolios to relevant geographic representatives (for example, the Fisheries ministry is traditionally given to an MP from one of Canada's coastal provinces), and increasingly diversity of gender and ethnicity. A survival analysis of cabinet entry in Canada found that gender, legal experience, education, age, and past ministerial tenure were major predictors of selection. (Kerby 2009) Cabinets in Canada have become increasingly professionalized and technocratic, a trend also observed in other developed democracies (Lammers and Nyomarkay 1982; Pekkanen, Nyblade, and Krauss 2013; Berlinski, Dewan, Dowding, and Subrahmanyam 2009)

New Zealand is a Westminster parliamentary democracy established in 1852. Historically it featured a lower house with fewer than 100 members and an SMP electoral system with reserved electorates for voters who self-identified as indigenous Maori. Dissatisfaction with the two major political parties, National and Labour, and representation concerns driven by majority governments with weak plurality electoral support, led New Zealand to enact electoral reform beginning with the 1996 national election. The current electoral system is closed-list mixed-member proportional and consists of approximately 121 seats including 71 electorate seats. The effective number of political parties varied from 1.76 to 2.16 before MMP and from 2.78 to 3.76 after MMP. Each government post-MMP has featured either supply agreements or coalitions with minor parties. Elections in New Zealand are extremely low-expense, with major party candidates spending in the order of \$20,000 per three-year election.

In principle, Labour governments in New Zealand allow their entire caucus to vote on cabinet membership, but in practice Prime Ministers are allowed considerable leeway to expand or shrink the size of cabinet and allocate portfolios among those selected. (Alley 1989; Wood 1989) National governments have no such policy, and the PM retains the ability to appoint at will. Cabinets typically include representation from both of New Zealand's islands and from urban and rural areas. The Minister of Maori Affairs position is often filled by a Maori legislator, and increasingly the representation of women in cabinet has been a relevant factor. (Curtin 2015)

Both countries share a variety of cultural and institutional factors: both feature high levels of parliamentary discipline and practice cabinet collective and individual responsibility. Expulsion from cabinet, caucus, and denial of renomination are the primary punishments available to enforce party discipline. Neither country has a formal primary election mechanism, instead delegating candidate selection to district-level parties, who by custom do not remove incumbents for reasons other than non-compliance with party leadership or scandal. (Levine 1979; Wood 1989) Finally, both countries practice ``government by cabinet'', with inert backbenches and powerful cabinets. (Henderson 1989)

@@

\section*{Data}
The dataset I make use of is novel. Existing election datasets suffer from limitations that render them inappropriate for use here: much of the data is aggregated above the constituency level; constituency level data is often missing names and characteristics of individual candidates necessary to make the data into a panel; district and candidate names vary across elections; and data typically does not include cabinet membership covariates.

As a result, I created a dataset containing full, constituency-level electoral data for all candidates and legislators in Canada from 1867 forward, and New Zealand from 1945 forward.\footnote{Canadian cabinet ministers before 1931 were expected to resign from the House of Commons upon being appointed and then run in by-elections to regain the seat they resigned. To avoid the complexity of this feature of the data and a party system transition before and during World War II, I discard data before the 1945 federal election.} Canadian data was obtained by writing a series of web scrapers and harvested from Canadian Library of Parliament sources. New Zealand data was assembled using a combination of previously published volumes\footnote{Wood 1996; Norton 1988}, paper returns supplied by the NZ Electoral Commission and digitized for the first time for this project, and web scrapers for data from 1996 onwards.

After combining and cleaning the data, the next challenge was being able to connect legislators across elections. To deal with internal inconsistencies in the reporting format of district names, party names, and candidate names, I wrote a text analysis program that uses candidate metadata to match candidates across elections to assemble the panel. Clear candidate matches were automatically identified, while unclear matches prompt the user to make a manual coding decision which is saved for later replication.

My dataset contains all candidates who stood for election in a district, but a small number of New Zealand MPs have been elected from the party list without contesting an electorate, including some cabinet ministers. These MPs are not represented in the dataset as they did not stand for election individually.\footnote{One example: Deputy Prime Minister Bill English, a long-time electorate incumbent who switched to list-only despite huge electorate majorities to minimize his travel commitment and time away from family. There is no electoral benefit to being list-only.}

@@

The primary limitation when testing this hypothesis is severe endogeneity. The ``gold standard'' to tackle this problem would be to experimentally appoint cabinet ministers through random assignment and measure causal effects. Reality is not so kind: Parties select cabinet ministers to satisfy party goals including rewarding key constituencies, demographics, and policy demanders within the party and coalition partners. Moreover, parties have access to private signals about incumbent quality learned during the recruitment, vetting, and campaigning process that are not fully accounted for in vote share at the time of their first election. These latent candidate qualities may become obvious during the process of legislating and materialize in future results, but are not strictly speaking a cabinet effect. Koop and Bittner (2011) argue that ``star candidates'' are much more likely to be appointed to cabinet. In light of these concerns, the claims made here cannot be said to be causal: effects presented are associations, suggestive of a relationship but not decisive.

@@

\setstretch{1.5}
\begin{itemize}
\item \textit{VotePct}: The vote share of the candidate.
\item \textit{CabinetNow}: The candidate was a member of the cabinet during the preceding term. I exclude ministers outside cabinet, junior ministers, ministers without portfolio, and other quasi-cabinet roles but include deputy Prime Ministers and Attorneys General.\footnote{I do not code ministerial exit from cabinet or role-switching mid-term, so in principle some of the \textit{CabinetNow} incumbents have been removed from cabinet before the election. Evidence suggests that ministerial exit is predicted by poor performance (Berlinkski, Dewan, and Dowding 2010) or scandal, (Dewan and Myatt 2007; Dewan and Dowding 2005) so this modeling decision should cause effect estimates to be biased conservatively as these ``false positive'' ministers drag the overall effect down.}
\item \textit{CabinetImportant}: The candidate was an ``important'' minister of the cabinet during the preceding term. ``Importance'' is subjective and here includes Defence, Health, Justice, Finance, Public Works, Revenue, and Foreign ministries. This variable exists to capture heterogeneity in the cabinet effect; senior ministers are presumed to have more discretion with respect to particularistic spending, more control over policy outcomes, and greater capacity to cultivate a personal vote through media exposure and credit claiming.
\item \textit{CabinetPM}: Dummy variable for the Prime Minister, regardless of whether they were an incumbent member of the legislature. The inclusion of a control for Prime Ministers is intended to absorb the positive electoral effects expected for a sitting Prime Minister to avoid their misattribution to cabinet membership.
\item \textit{Incumbent}: The candidate is an incumbent member of the legislature. I treat candidates as incumbent even if redistricting occurs, or if the candidate was previously elected by party list and now contests an electorate. Incumbency is known to be a a strong predictor of re-election.
\item \textit{TermsServed}: The total unbroken number of terms served by a candidate. Tenure does not reset if an incumbent resigns and runs in a by-election, but does reset if a candidate loses or resigns and later returns to the legislature. I treat by-elected members as serving a full time. Tenure is present in the model to capture the accumulated effects of developing a personal vote (or, at least, a reputation of competency) in an incumbent's district over time.
\item \textit{Age}: The age of the candidate at the time of the election. Data is only available for a subset of Canadian candidates who were elected to office at least once. Age was gathered to allow for a proxy for signals of candidate quality including occupation, past electoral and career experience.
\item \textit{Elected}: Binary indicator for election outcome.
\item \textit{Party}: Party label the member stood for election under. I recode minor parties or splinter factions to be independent.
\item \textit{PartyInGovt}: Is the candidate running for the party currently in power? In New Zealand post-1996, I include junior coalition partner parties but not parties in supply agreements with the governing party. This variable was constructed in order to account for the ``electoral costs of governing''.
\end{itemize}
\setstretch{2}

@@

The effect's appearance in Canada contrasts its absence in New Zealand. Effect estimates in the main model are small (1.3\%) with confidence intervals spanning substantively unimportant territory. Martin's model reveals a small effect of 2.1\%, but is likely picking up correlation between tenure and cabinet selection due to the excluded tenure variable. Additional specifications are inconsistent as to whether the effect size is substantively interesting. Disaggregating by decade reveals no consistency. Further analysis as to why New Zealand may behave differently than Ireland or Canada is left until after presentation of results.

@@

Cabinet ministers are more popular, but not necessarily because they are cabinet ministers. In fact, there is evidence the causality is exactly reversed: A cross-sectional logistic regression of probability being selected into the Canadian Cabinet based on age, tenure, and previous performance suggests that previous performance matters a great deal. In the modal case, a 50 year old sophomore MP of the governing party is much 50\% more likely to be appointed a cabinet minister if she comes from a 70\% blowout win than a 40\% plurality win.  Controlling for past performance can help establish temporal ordering.

@@

\setstretch{1.25}
\begin{eqnarray*}
	\Delta y_i &=& \beta X_i + \epsilon_i =\\
	y_{i,t} - y_{i,t-1} &=& \beta X_i + \epsilon_i = \\
	y_{i,t} &=& \beta X_i + y_{i,t-1} + \epsilon_i
\end{eqnarray*}
\setstretch{2}
The assumption about the relationship between previous performance and current performance is explicit. Incumbents begin with their previous vote share, and the model fit determines how the covariates are associated with deviations from that share.\footnote{This is conceptually difference from a true ``first differences'' estimator, which also differences out predictor covariates.} A second choice to control for past performance is a lagged dependent variable. This is a weaker assumption; in this model, past performance is merely associated with present vote share, with the strength of the relationship permitted to vary:

@@

\setstretch{2} As expected, endogeneity stemming from the relationship between popularity and cabinet selection in the previous term drives some of the previous effect. Controlling for past performance, the coefficient estimate of \textit{CabinetNow} (as well as tenure and party) decreases markedly in Canada. In the primary, controlling-for-past-performance model, the effect size is reduced to 1.2\% on average (with the confidence interval permitting interpretations of anything from essentially no effect to an effect slightly smaller than initial cross-sectional estimates). The $\Delta VotePct$ model effects are even smaller still at 0.8\% and include the possibility of no effect.\footnote{A third possibility to control for past performance is to introduce individual fixed effects. This model also results in a smaller estimate of approximately 0.7\% effect for cabinet service with a confidence interval including no effect.} In New Zealand, meanwhile, the effect is negative in all three specifications, although modest in each.

Roughly three quarters of the original dataset are dropped to run these models, causing a loss of power---but it is clear that the effect magnitude is reduced in Canada and entirely compromised in New Zealand. Serial autocorrelation in the error structure induced by this strategy is another potential threat to estimate validity. Disaggregation of the CA Control model by decade reveals further inconsistency in the result, with striking effects in the 1980s and 1990s, weaker effects in the 1950s, and near-zero or weakly negative effects in other decades.

@@

Again, the effect is visible and substantively relevant, although modest in size, in Canada. Logistic regression coefficients are not clearly interpretable, but when transformed into odds ratios and predicted probabilities, they become clearer. Using bootstrap resampling to estimate the effect of cabinet membership\footnote{\noindent 1000 samples via nonparametric bootstrap, calculating predicted probabilities for modal case in and out of cabinet, and taking difference to recover effect and CI.}, a two-term incumbent Liberal backbencher in a Liberal government has a 73\% (71\% - 76\%) probability to be re-elected while a cabinet minister with the same characteristics has an 81\% (75\% - 86\%) probability of re-election. A 95\% empirical confidence interval of the effect spans 1.2\%-12.5\%, centered at 7.5\%. As with the above example, the effect is about half the size of an additional term of tenure: modest, but real. The result remains invariant to alternate model specifications tested, including consistent positive effects (some with too little power to rule out no effect) across the disaggregation by decade.

@@

The assumption underpinning the Cox PH model is proportionality; in other words, that covariates affect the hazard rate in a constant way across the time variable. A test of this assumption using Schoenfeld residuals finds insufficient evidence to reject the null hypothesis (proportional hazards) for the \textit{CabinetNow} variable and validates the appropriateness of the Cox PH model.
